Estimating autocorrelation from multiple Markov chains

### 2020-06-03 Paper/main.tex
Crisis diverted

\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{commath}
\usepackage{enumerate}
\newtheorem{theorem}{Theorem}
\usepackage{amsmath}
\usepackage{natbib}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\graphicspath{ {./plots/} }
\DeclareMathOperator{\Tr}{Tr}
\newcommand{\dd}[1]{\mathrm{d}#1}
\newcommand{\E}{\mathbb{E}}
\newcommand{\Var}{\text{Var}}
\newcommand{\X}{\mathcal{X}}
\newcommand{\B}{\mathcal{B}}
% These three packages are from the American Mathematical Society and includes all of the important symbols and operations 
\usepackage{fullpage}
\usepackage{comment}
% By default, an article has some vary large margins to fit the smaller page format.  This allows us to use more standard margins.

\newtheorem{ass}{Assumption}
\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}
\newtheorem{remark}{Remark}
\setlength{\parskip}{0em}
% This gives us a full line break when we write a new paragraph
\usepackage[mathscr]{euscript}
\usepackage{tikz}
\newcommand*\circled[1]{\tikz[baseline=(char.base)]{
            \node[shape=circle,draw,inner sep=2pt] (char) {#1};}}
\newcommand{\indep}{\perp \!\!\! \perp}
\title{Estimating autocorrelation from multiple Markov chains}

\author{
  Medha Agarwal \\
  \textit{Department of Mathematics and Statistics}\\
  \textit{IIT Kanpur}\\
  \texttt{medhaaga@iitk.ac.in} \\\\
  %% examples of more authors
 Dootika Vats \\
  \textit{Department of Mathematics and Statistics}\\
  \textit{IIT Kanpur}\\
  \texttt{dootika@iitk.ac.in} \\
}

@@

\begin{abstract}
    This is best abstract in the world.\\.\\.\\.\\.\\.
\end{abstract}

@@

Throughout this paper, we will deal with the function $g(X) = X$. We denote the sample mean of $s^{th}$ Markov chain by $\overline{X}_s$ and the overall mean by $\overline{\overline{X}} = \sum_{s = 1}^{m}\overline{X}_s/m$. We are interested in Monte Carlo error, i.e. $\overline{\overline{X}} - \mu$ to assess the reliability of simulations. The sampling distribution for Monte Carlo error of $s^{th}$ chain is available via Markov chain CLT (\cite{jones2004markov}) if there exists a positive definite matrix $\Sigma$ such that

\[
\sqrt{n}(\overline{X}_s-\mu) \xrightarrow{d} N(0,\Sigma)
\]

 With $m$ i.i.d. samples of $\overline{X}_s$, $s \in \{1,..., m\}$, the MCSE is given by $\Sigma/mn$. We are interested in consistent estimates of $\Sigma$ for two reasons. Firstly, to construct asymptotically valid confidence intervals and secondly, to determine when to stop the simulations. For this purpose, we use the class of Multivariate Spectral Variance Estimators (MSVE) wherein we use R-ACF to estimate the autocovariance. We call this replicated spectral variance (RSV) estimator (see \cite{arg:and:2006} for replicated batch means). \\

@@

\[
\hat{\Gamma}_{RAC,s}(k) = \dfrac{1}{n}\sum_{t=1}^{n-k}(X_{st}-\overline{\overline{X}})(X_{s(t+k)}-\overline{\overline{X}})^T
\]

@@

\begin{figure}
\begin{subfigure}{\textwidth}
  \centering
  \includegraphics[width=.8\linewidth]{var-acf_5e2.pdf}
  \caption{$n = 500$}
  \label{subfig:acf-500}
\end{subfigure}\\
\begin{subfigure}{\textwidth}
  \centering
  \includegraphics[width=.8\linewidth]{var-acf_5e4.pdf}
  \caption{$n = 5*10^4$}
  \label{subfig:acf-5e4}
\end{subfigure}
\caption{Autocorrelation plot for ACF and R-ACF for first chain out of two parallel Markov chains. First column corresponds to ACF calculated using arithmetic mean of first Markov chain and the second column corresponds to the one calculated using global mean of two Markov chains.(a): $n = 500$, not converged yet; (b): $n = 5 \times 10^4$, converged}
\label{fig:var_acf_ccf}
\end{figure}

@@

By proposition 1 in \cite{song1995optimal} 
\begin{equation} \label{eq:song}
\Var(\overline{X}) = \dfrac{\Sigma}{n} + \dfrac{\Phi}{n^2} + o(n^{-2})
\end{equation}

As a consequence, if $\Var(\bar{X})$ is finite, then $\Var(\bar{X}) \to 0 \textrm{ as } n \to \infty$. We can break $\hat{\Gamma}_{RAC}$ into four parts for all $k \geq 1$ as:
 \[
 \hat{\Gamma}_{RAC} =  \dfrac{1}{m}\sum_{s=1}^{m}\hat{\Gamma}_s(k) - \dfrac{1}{mn}\sum_{s=1}^{m}\sum_{t=1}^{|k|}A_{st}^T - \dfrac{1}{mn}\sum_{s=1}^{m}\sum_{t=n-|k|+1}^{n}A_{st} + \dfrac{n-|k|}{mn}\sum_{s=1}^{m}(\overline{X_s}-\overline{\overline{X}})(\overline{X_s}-\overline{\overline{X}})^T
 \]

where $A_{st} = (X_{st}-\overline{X}_s)(\overline{X}_s - \overline{\overline{X}})^T$\\
The expectation value of the first term is given by equation \ref{eq:priestly} and \ref{eq:song} while the other three terms contribute $o(1)$ terms. The complete proof evaluating the expectation values of last three terms can be found in Appendix subsection \ref{appendix:bias}.
\end{proof}

@@

\begin{corollary} \label{cor:lag0_expectation}, 
\[
\mathbb{E}[\hat{\Gamma}_{RAC}(0)] = \Gamma(0) + \dfrac{\Sigma}{mn} + \dfrac{\Phi}{mn^2} + o(n^{-2})
\]
\end{corollary}

@@

\[
\hat{\Gamma}(k) = \dfrac{1}{n}\sum_{t=1}^{n-\abs{k}}(X_t - \overline{X})(X_{t + \abs{k}} - \overline{X})^T
\]

@@

\[
\hat{\Sigma}_{SV} &= \sum_{k=-b_n+1}^{b_n-1}w_n\left(\dfrac{k}{b_n}\right)\hat{\Gamma}_n(k)
\]

where $w_n(.)$ is the lag window and $b_n$ is the truncation point.\\

 Consider $m$ Markov chains where $\overline{X}_s$, $\hat{\Gamma}_s(k)$ and $\hat{\Sigma}_{SV, s}$ are the sample mean, sample lag-k autocovariance and spectral variance estimators respectively for $s^{th}$ Markov chain, $s\in {1,...,m}$. We will consider two estimators for spectral variance calculations using multiple Markov chains.

@@

\begin{align*}
    \hat{\Sigma}_{SV,s} &= \sum_{k=-b_n+1}^{b_n-1}w_n\left(\dfrac{k}{b_n}\right)\hat{\Gamma}_s(k)\\
    \hat{\Gamma}_s(k) &= \dfrac{1}{n}\sum_{t=1}^{n-k}(X_{st}-\overline{X_s})(X_{s(t+k)}-\overline{X_s})^T
\end{align*}

@@

Now we introduce the Replicated Spectral Variance Estimator (RSV). $\hat{\Gamma}_{RAC,s}(k)$ is the sample replicated lag-k autocovariance estimator for the $s^{th}$ MC using overall mean $(\overline{\overline{X}})$ instead of $\overline{X}_s$ for centering. We construct an RSV estimator $\hat{\Sigma}_{RSV} $ using this replicated version of autocovariance estimator. Consider,

\begin{align*}
    \hat{\Sigma}_{RSV,s} &= \sum_{k=-b_n+1}^{b_n-1}w_n\left(\dfrac{k}{b_n}\right)\hat{\Gamma}_{RAC,s}(k)\\
\end{align*}
$$\hat{\Sigma}_{RSV} &=  \dfrac{1}{m}\sum_{s=1}^{m}\hat{\Sigma}_{RSV,s}$$

@@

\begin{align*}
    \tilde{\Sigma}_s &= \sum_{k=-b_n+1}^{b_n-1}w_n\left(\dfrac{k}{b_n}\right)\tilde{\Gamma}_s(k)\\
    \tilde{\Gamma}_s(k) &= \dfrac{1}{n}\sum_{t=1}^{n-|k|}(X_{st}-\mu)(X_{s(t+k)}-\mu)^T
\end{align*}

@@

\bigskip
\begin{theorem}
\label{th:consistency}
 Let the Assumptions \ref{ass:sip},\ref{ass:sve_consis} hold and $\dfrac{b_n \log \log n}{n} \to 0 \textrm{ as } n \to \infty$, then $\hat{\Sigma}_{RSV} \to \Sigma$ with probability 1, as $n \to \infty$.
\end{theorem}

@@

\[
\hat{\Sigma}_{RSV}^{ij} - \tilde{\Sigma}^{ij} = g_1(n)D^2 + g_2(n)D + g_3(n)
\]
where $D$ is the finite random variable associated with SIP. We will show that\\

\begin{align*}
    g_1(n) &= (4+C_1)\dfrac{b_n \psi^2(n)}{n^2} - 4\dfrac{\psi^2(n)}{n^2}\\
    g_2(n) &= 2\sqrt{2}\|L\|p^{1/2}(1+\epsilon)\left[(4+C_1)\dfrac{b_n\psi(n)\sqrt{n\log \log n}}{n^2} - 4\dfrac{\psi(n)\sqrt{n\log \log n}}{n^2}\right]\\
    g_3(n) &= \|L\|^2 p (1+\epsilon)^2\left[(4+C_1)\dfrac{b_n \log\log n}{n} - 4 \dfrac{\log \log n}{n}\right]
\end{align*}

@@

We are interested in the finding the bias for RSV estimator as a function of $n$. The proof below applies to all the important range of lag windows $w_n(x)$; where $w_n(x)$ is a continuous, even fnction with $w_n(0)=1, \abs{w_n(x)} < c$, and $\int_{-\infty}^{\infty}w_n^2(x)dx < \infty$. Define

\[
W_n = \dfrac{1}{2\pi}\sum_{k=-b_n+1}^{b_n-1}w_n\left(\dfrac{k}{b_n}\right)
\]

@@

\begin{ass} \label{ass:bias}
    Let $\Gamma_s(k)$ be the lag-k autocovariance for $s^{th}$ Markov chain and $w_n(x)$ be the lag window. We assume that there exists a $q \geq 0$ such that
    \begin{enumerate} [a.]
        \item $\sum\limits_{-\infty}^{\infty}\abs{k}^q \|\Gamma(k)\| < \infty$
        \item $\lim _{x \to 0}\dfrac{1 - w_n(x)}{\abs{x}^q} = k_q < \infty$
        \item $\dfrac{b_n^q}{n} \to 0$ as $n \to \infty$
    \end{enumerate}

@@

\begin{theorem}\label{th:rsv_bias}
The limiting bias of replicated spectral variance is given by:
\[
 \lim_{n \to \infty}b_n^q\mathbb{E}[\hat{\Sigma}_{RSV} - \Sigma] = -k_q\Phi^{(q)}
 \]
\end{theorem}

\begin{proof}

\[
\mathbb{E}[\hat{\Sigma}_{RSV} - \Sigma] = \sum_{k=-b_n+1}^{b_n-1} w\left(\dfrac{|k|}{b_n}\right)\mathbb{E}[\hat{\Gamma}_{RAC}(k)]-\sum_{k=-\infty}^{\infty}\Gamma(k)
\]

@@

\begin{theorem} \label{th:rsv_variance}
 If Assumption \ref{ass:variance_cal} hold, $\dfrac{n}{b_n}\Var\hat{\Sigma}_{RSV}^{ij} = [\Sigma_{ii}\Sigma_{jj} + \Sigma_{ij}^2]\int_{-\infty}^{\infty}w(x)^2dx  + o(1)$.
\end{theorem}

@@

\[
\textrm{Volume of confidence region}^{1/p} + n^{-1} < \epsilon \abs{\Lambda_n}^{1/2p} 
\]

where $\epsilon$ is the tolerance level, $\abs{.}$ denotes the determinant and $\Lambda_n$ is the sample covariance matrix given by

\[
\dfrac{1}{mn-1}\sum_{s=1}^{m}\sum_{t=1}^{n}(X_{st} - \overline{\overline{X}})(X_{st} - \overline{\overline{X}})^T
\]

The determinant of Monte Carlo standard error is called generalized variance in  \cite{wilks1932certain} and gives a metric for volume of confidence interval. \cite{vats2019multivariate} showed that if the estimator for $\Sigma$ is strongly consistent, the stopping rule is asymptotically valid, in that the confidence regions created at termination have the right coverage probability as $\epsilon \to 0$. Since $\hat{\Sigma}_{RSV}$ is a strongly consistent estimator of $\Sigma$, w

@@

\[
ESS = mn\left(\dfrac{\abs{\Lambda_{mn}}}{\abs{\hat{\Sigma}_{RSV}}}\right)^{1/p}
\]

@@

We examine the vector autoregressive process of order 1 (VAR(1)) where the true autocovariance in known in closed form. We have already illustrated the better performance of R-ACF over traditional ACF on a slowly mixing VAR(1) process in Figure \ref{fig:var_acf_ccf}. Here we will examine the performance of RSV estimator on the same process. Consider a p-dimensional VAR(1) process $\{X_t\}_{t \geq 1}$ such that

\[
X_t = \Phi X_{t-1} + \epsilon_t
\]

where $X_t \in \mathbb{R}^p$, $\Phi $ is a $p \times p $ matrix, $ \epsilon_t \overset{i.i.d.}{\sim} N(0, \Omega)$, and $\Omega$ is a positive definite $p \times p$ matrix. The invariant distribution for this Markov chain is $N(0, \Psi)$ where $\Vec{\Psi} = (I_{p^2} - \Phi \otimes \Phi)\Vec{\Omega}$. The lag-k autocovariance can be calculated easily for $k >0$ as

\begin{align*}
    \Gamma(k) &= \Phi^k\Psi\\
    \Gamma(-k) &= \Psi(\Phi^T)^k
\end{align*}

The Markov chain is geometrically ergodic when the spectral norm of $\Phi$ is less than 1 (\cite{10.2307/1427459}). The CLT holds for the invariant distribution, therefore, $\Sigma$ exists and is given by

\[
\Sigma = (1 - \Phi)^{-1}\Psi + \Psi(1 - \Phi^T)^{-1} - \Psi
\]

@@

\begin{figure}[h]
    \centering
    \begin{subfigure}[h]{0.4\textwidth}
         \centering
         \includegraphics[width=\textwidth]{var-sp_5e2.pdf}
         \caption{$n = 500$}
         \label{subfig:var-sp_500}
     \end{subfigure}
     \begin{subfigure}[h]{0.4\textwidth}
         \centering
         \includegraphics[width=\textwidth]{var-sp_5e4.pdf}
         \caption{$n = 5 \times 10^4$}
         \label{subfig:var-sp_50000}
     \end{subfigure}
    \caption{Scatter plot of five Markov chains for different sample sizes.}
    \label{fig:var-sp}
\end{figure}

@@

We will use a bivariate bi-modal distribution introduced by \cite{gelman1991note} which has Gaussian conditional distributions in both directions. This allows us to sample parallel Markov chains using the Gibbs sampler. Let $x$ and $y$ be two random variable that are jointly distributed as

\[
f(x, y) \propto \exp\left(-\dfrac{1}{2}[Ax^2y^2 + x^2 + y^2 -2Bxy  -2C_1x - 2C_2y]\right)
\]

The conditional distribution of $x$ given $y$ and vice versa is a normal distribution given by:

\begin{align*}
    x_1 \mid x_2 &\sim N\left(\dfrac{Bx_2 + C_1}{Ax_2^2 + 1}, \dfrac{1}{Ax_2^2 + 1}\right)\\
    x_2 \mid x_1 &\sim N\left(\dfrac{Bx_1 + C_2}{Ax_1^2 + 1}, \dfrac{1}{Ax_1^2 + 1}\right)
\end{align*}

@@

In Figure \ref{fig:boom-sp}, we demonstrate the "sticky" nature of the Markov chains. For the first thousand samples, both the chains are oblivious of the existence of another mode. By ten thousand samples, both the Markov chains have explored the two modes. This property will affect the single chain estimators like ACF. In Figure \ref{subfig:boom-acf_1e3}, the ACF is severely underestimated because the chain-1 has not jumped to the other mode. Whereas, in Figure \ref{subfig:boom-acf_1e4}, both ACF and R-ACF give almost similar results indicating that both modes have been explored by chain-1.

@@

\begin{table}[h]
\centering
    \begin{tabular}{p{1cm}|p{2cm}|p{2cm}|p{2cm}|p{2cm}}
    \toprule
        n & \multicolumn{2}{|c|}{$m = 2$} & \multicolumn{2}{|c|}{$m=5$}\\
        \hline
        & ASV & RSV & ASV & RSV\\
        \hline
        $10^3$ & 0.612 & 0.689 & 0.602 & 0.753   \\
        $2\cdot10^3$ & 0.693 & 0.751 & 0.735 & 0.827 \\
        $5 \cdot 10^3$  & 0.826 & 0.854 & 0.847 & 0.880 \\
        $10^4$  & 0.862 & 0.868 & 0.884 & 0.907 \\
        $2 \cdot 10^4$ & 0.899 & 0.906 & 0.922 & 0.934 &  &  \\
    \bottomrule
    \end{tabular}
    \caption{Coverage probabilities for parameter values $A = 1,\; B = 3,\; C = 8$. }
    \label{table:boom-coverage_1_3_8}
\end{table}

@@

\begin{lemma} \label{lemma:consis_1}
  Let the strong invariance principle hold. If $\dfrac{\psi(n)}{n} \to 0$ as $n \to \infty$, then $\|\overline\overline{{X}} - \mu\|_{\infty}, \|\overline{X}_s - \mu\|_{\infty} \xrightarrow[]{a.s.} 0$ as $n \to \infty$ for all $s \in \{1,..., m\}$\\
  \end{lemma}

>>>>

\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{commath}
\usepackage{enumerate}
\usepackage{amsmath}
\usepackage{natbib}
\usepackage{graphicx}
\usepackage{caption,setspace}
\usepackage{subcaption}
\graphicspath{ {./plots/} }
\DeclareMathOperator{\Tr}{Tr}
\newcommand{\dd}[1]{\mathrm{d}#1}
\newcommand{\E}{\mathbb{E}}
\newcommand{\Var}{\text{Var}}
\newcommand{\X}{\mathcal{X}}
\newcommand{\B}{\mathcal{B}}
% These three packages are from the American Mathematical Society and includes all of the important symbols and operations 
\usepackage{fullpage}
\usepackage{comment}
% By default, an article has some vary large margins to fit the smaller page format.  This allows us to use more standard margins.
\allowdisplaybreaks
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}

\theoremstyle{remark}
\newtheorem{ass}{Assumption}
\newtheorem{remark}{Remark}
\setlength{\parskip}{0em}
% This gives us a full line break when we write a new paragraph
% \usepackage[mathscr]{euscript}
\usepackage{tikz}
\newcommand*\circled[1]{\tikz[baseline=(char.base)]{
            \node[shape=circle,draw,inner sep=2pt] (char) {#1};}}
% \newcommand{\indep}{\perp \!\!\! \perp}

\title{Estimating autocorrelation from multiple Markov chains}

\author{
  Medha Agarwal \\
  Department of Mathematics and Statistics\\
  IIT Kanpur\\
  \texttt{medhaaga@iitk.ac.in} \and
  %% examples of more authors
 Dootika Vats \\
  Department of Mathematics and Statistics\\
  IIT Kanpur\\
  \texttt{dootika@iitk.ac.in}
}

@@

\onehalfspacing
\begin{abstract}
    This is best abstract in the world.\\.\\.\\.\\.\\.
\end{abstract}

@@

Throughout this paper, we will deal with the function $g(X) = X$. We denote the sample mean of $s^{th}$ Markov chain by $\bar{X}_s$ and the overall mean by $\bar{\bar{X}} = \sum_{s = 1}^{m}\bar{X}_s/m$. We are interested in Monte Carlo error, i.e. $\bar{\bar{X}} - \mu$ to assess the reliability of simulations. The sampling distribution for Monte Carlo error of $s^{th}$ chain is available via Markov chain CLT (\cite{jones2004markov}) if there exists a positive definite matrix $\Sigma$ such that

\[
\sqrt{n}(\bar{X}_s-\mu) \xrightarrow{d} N(0,\Sigma)
\]

 With $m$ i.i.d. samples of $\bar{X}_s$, $s \in \{1,..., m\}$, the MCSE is given by $\Sigma/mn$. We are interested in consistent estimates of $\Sigma$ for two reasons. Firstly, to construct asymptotically valid confidence intervals and secondly, to determine when to stop the simulations. For this purpose, we use the class of Multivariate Spectral Variance Estimators (MSVE) wherein we use R-ACF to estimate the autocovariance. We call this replicated spectral variance (RSV) estimator (see \cite{arg:and:2006} for replicated batch means). \\

@@

\[
\hat{\Gamma}_{RAC,s}(k) = \dfrac{1}{n}\sum_{t=1}^{n-k}(X_{st}-\bar{\bar{X}})(X_{s(t+k)}-\bar{\bar{X}})^T
\]

@@

\begin{figure}
\begin{subfigure}{\textwidth}
  \centering
  \includegraphics[width=.8\linewidth]{plots/acf-500.pdf}
  \caption{$n = 500$}
  \label{subfig:acf-500}
\end{subfigure}\\
\begin{subfigure}{\textwidth}
  \centering
  \includegraphics[width=.8\linewidth]{plots/acf-5e4.pdf}
  \caption{$n = 5*10^4$}
  \label{subfig:acf-5e4}
\end{subfigure}
\caption{Autocorrelation plot for ACF and R-ACF for first chain out of two parallel Markov chains. First column corresponds to ACF calculated using arithmetic mean of first Markov chain and the second column corresponds to the one calculated using global mean of two Markov chains.(a): $n = 500$, not converged yet; (b): $n = 5 \times 10^4$, converged}
\label{fig:var_acf_ccf}
\end{figure}

@@

By proposition 1 in \cite{song1995optimal} 
\begin{equation} \label{eq:song}
\Var(\bar{X}) = \dfrac{\Sigma}{n} + \dfrac{\Phi}{n^2} + o(n^{-2})
\end{equation}

As a consequence, if $\Var(\bar{X})$ is finite, then $\Var(\bar{X}) \to 0 \textrm{ as } n \to \infty$. We can break $\hat{\Gamma}_{RAC}$ into four parts for all $k \geq 1$ as:
 \[
 \hat{\Gamma}_{RAC} =  \dfrac{1}{m}\sum_{s=1}^{m}\hat{\Gamma}_s(k) - \dfrac{1}{mn}\sum_{s=1}^{m}\sum_{t=1}^{|k|}A_{st}^T - \dfrac{1}{mn}\sum_{s=1}^{m}\sum_{t=n-|k|+1}^{n}A_{st} + \dfrac{n-|k|}{mn}\sum_{s=1}^{m}(\bar{X}_s - \bar{\bar{X}})(\bar{X}_s - \bar{\bar{X}})^T\,,
 \]
%
where $A_{st} = (X_{st}-\bar{X}_s)(\bar{X}_s - \bar{\bar{X}})^T$\\
The expectation value of the first term is given by equation \ref{eq:priestly} and \ref{eq:song} while the other three terms contribute $o(1)$ terms. The complete proof evaluating the expectation values of last three terms can be found in Appendix subsection \ref{appendix:bias}.
\end{proof}

@@

\begin{corollary} \label{cor:lag0_expectation}
\[
\mathbb{E}[\hat{\Gamma}_{RAC}(0)] = \Gamma(0) + \dfrac{\Sigma}{mn} + \dfrac{\Phi}{mn^2} + o(n^{-2})
\]
\end{corollary}

@@

\[
\hat{\Gamma}(k) = \dfrac{1}{n}\sum_{t=1}^{n-\abs{k}}(X_t - \bar{X})(X_{t + \abs{k}} - \bar{X})^T
\]

@@

\[
\hat{\Sigma}_{SV} = \sum_{k=-b_n+1}^{b_n-1}w\left(\dfrac{k}{b_n}\right)\hat{\Gamma}_n(k)
\]

where $w(.)$ is the lag window and $b_n$ is the truncation point.\\

 Consider $m$ Markov chains where $\bar{X}_s$, $\hat{\Gamma}_s(k)$ and $\hat{\Sigma}_{SV, s}$ are the sample mean, sample lag-k autocovariance and spectral variance estimators respectively for $s^{th}$ Markov chain, $s\in {1,...,m}$. We will consider two estimators for spectral variance calculations using multiple Markov chains.

@@

\begin{align*}
    \hat{\Sigma}_{SV,s} &= \sum_{k=-b_n+1}^{b_n-1}w\left(\dfrac{k}{b_n}\right)\hat{\Gamma}_s(k)\\
    \hat{\Gamma}_s(k) &= \dfrac{1}{n}\sum_{t=1}^{n-k}(X_{st}-\overline{X_s})(X_{s(t+k)}-\overline{X_s})^T
\end{align*}

@@

Now we introduce the Replicated Spectral Variance Estimator (RSV). $\hat{\Gamma}_{RAC,s}(k)$ is the sample replicated lag-k autocovariance estimator for the $s^{th}$ MC using overall mean $(\bar{\bar{X}})$ instead of $\bar{X}_s$ for centering. We construct an RSV estimator $\hat{\Sigma}_{RSV} $ using this replicated version of autocovariance estimator. Consider,

\begin{align*}
    \hat{\Sigma}_{RSV,s} &= \sum_{k=-b_n+1}^{b_n-1}w\left(\dfrac{k}{b_n}\right)\hat{\Gamma}_{RAC,s}(k)
\end{align*}
\[
\hat{\Sigma}_{RSV} =  \dfrac{1}{m}\sum_{s=1}^{m}\hat{\Sigma}_{RSV,s}
\]

@@

\begin{align*}
    \tilde{\Sigma}_s &= \sum_{k=-b_n+1}^{b_n-1}w\left(\dfrac{k}{b_n}\right)\tilde{\Gamma}_s(k)\\
    \tilde{\Gamma}_s(k) &= \dfrac{1}{n}\sum_{t=1}^{n-|k|}(X_{st}-\mu)(X_{s(t+k)}-\mu)^T
\end{align*}

@@

\bigskip
\begin{theorem}
\label{th:consistency}
 Let the Assumptions \ref{ass:sip},\ref{ass:sve_consis} hold and $n^{-1}{b_n \log \log n} \to 0 \textrm{ as } n \to \infty$, then $\hat{\Sigma}_{RSV} \to \Sigma$ with probability 1, as $n \to \infty$.
\end{theorem}

@@

\[
\hat{\Sigma}_{RSV}^{ij} - \tilde{\Sigma}^{ij} = g_1(n)D^2 + g_2(n)D + g_3(n)
\]
where $D$ is the finite random variable associated with SIP. We will show that
\begin{align*}
    g_1(n) &= (4+C_1)\dfrac{b_n \psi^2(n)}{n^2} - 4\dfrac{\psi^2(n)}{n^2}\\
    g_2(n) &= 2\sqrt{2}\|L\|p^{1/2}(1+\epsilon)\left[(4+C_1)\dfrac{b_n\psi(n)\sqrt{n\log \log n}}{n^2} - 4\dfrac{\psi(n)\sqrt{n\log \log n}}{n^2}\right]\\
    g_3(n) &= \|L\|^2 p (1+\epsilon)^2\left[(4+C_1)\dfrac{b_n \log\log n}{n} - 4 \dfrac{\log \log n}{n}\right]
\end{align*}

@@

We are interested in the finding the bias for RSV estimator as a function of $n$. The proof below applies to all the important range of lag windows $w(x)$; where $w(x)$ is a continuous, even fnction with $w(0)=1, \abs{w(x)} < c$, and $\int_{-\infty}^{\infty}w^2(x)dx < \infty$. Define

\[
W_n = \dfrac{1}{2\pi}\sum_{k=-b_n+1}^{b_n-1}w\left(\dfrac{k}{b_n}\right)
\]

@@

\begin{ass} \label{ass:bias}
    Let $\Gamma_s(k)$ be the lag-k autocovariance for $s^{th}$ Markov chain and $w(x)$ be the lag window. We assume that there exists a $q \geq 0$ such that
    \begin{enumerate} [a.]
        \item $\sum\limits_{-\infty}^{\infty}\abs{k}^q \|\Gamma(k)\| < \infty$
        \item $\lim _{x \to 0}\dfrac{1 - w(x)}{\abs{x}^q} = k_q < \infty$
        \item $\dfrac{b_n^q}{n} \to 0$ as $n \to \infty$
    \end{enumerate}

@@

\begin{theorem}\label{th:rsv_bias}
The limiting bias of replicated spectral variance is given by:
\[
 \lim_{n \to \infty}b_n^q\mathbb{E} \left[\hat{\Sigma}_{RSV} - \Sigma \right] = -k_q\Phi^{(q)}
 \]
\end{theorem}

\begin{proof}
\[
\mathbb{E} \left[\hat{\Sigma}_{RSV} - \Sigma \right] = \sum_{k=-b_n+1}^{b_n-1} w\left(\dfrac{|k|}{b_n}\right)\mathbb{E} \left[\hat{\Gamma}_{RAC}(k) \right] - \sum_{k=-\infty}^{\infty}\Gamma(k)
\]

@@

\begin{theorem} \label{th:rsv_variance}
 If Assumption \ref{ass:variance_cal} hold, $b_n^{-1}{n}\Var \left(\hat{\Sigma}_{RSV}^{ij} \right) = [\Sigma_{ii}\Sigma_{jj} + \Sigma_{ij}^2]\int_{-\infty}^{\infty}w(x)^2dx  + o(1)$.
\end{theorem}

@@

\[
\textrm{Volume of confidence region}^{1/p} + n^{-1} < \epsilon \abs{\Lambda_n}^{1/2p} 
\]
%
where $\epsilon$ is the tolerance level, $\abs{.}$ denotes the determinant and $\Lambda_n$ is the sample covariance matrix given by
%
\[
\dfrac{1}{mn-1}\sum_{s=1}^{m}\sum_{t=1}^{n}(X_{st} - \bar{\bar{X}})(X_{st} - \bar{\bar{X}})^T
\]
%
The determinant of Monte Carlo standard error is called generalized variance in  \cite{wilks1932certain} and gives a metric for volume of confidence interval. \cite{vats2019multivariate} showed that if the estimator for $\Sigma$ is strongly consistent, the stopping rule is asymptotically valid, in that the confidence regions created at termination have the right coverage probability as $\epsilon \to 0$. Since $\hat{\Sigma}_{RSV}$ is a strongly consistent estimator of $\Sigma$, w

@@

\[
\text{ESS} = mn\left(\dfrac{\abs{\Lambda_{mn}}}{\abs{\hat{\Sigma}_{RSV}}}\right)^{1/p}
\]

@@

We examine the vector autoregressive process of order 1 (VAR(1)) where the true autocovariance in known in closed form. We have already illustrated the better performance of R-ACF over traditional ACF on a slowly mixing VAR(1) process in Figure \ref{fig:var_acf_ccf}. Here we will examine the performance of RSV estimator on the same process. Consider a p-dimensional VAR(1) process $\{X_t\}_{t \geq 1}$ such that
%
\[
X_t = \Phi X_{t-1} + \epsilon_t
\]

where $X_t \in \mathbb{R}^p$, $\Phi $ is a $p \times p $ matrix, $ \epsilon_t \overset{i.i.d.}{\sim} N(0, \Omega)$, and $\Omega$ is a positive definite $p \times p$ matrix. The invariant distribution for this Markov chain is $N(0, \Psi)$ where $\Vec{\Psi} = (I_{p^2} - \Phi \otimes \Phi)\Vec{\Omega}$. The lag-k autocovariance can be calculated easily for $k >0$ as
%
\begin{align*}
    \Gamma(k) &= \Phi^k\Psi\\
    \Gamma(-k) &= \Psi(\Phi^T)^k
\end{align*}

The Markov chain is geometrically ergodic when the spectral norm of $\Phi$ is less than 1 (\cite{10.2307/1427459}). The CLT holds for the invariant distribution, therefore, $\Sigma$ exists and is given by
%
\[
\Sigma = (1 - \Phi)^{-1}\Psi + \Psi(1 - \Phi^T)^{-1} - \Psi
\]

@@

\begin{figure}[h]
    \centering
    \begin{subfigure}[h]{0.4\textwidth}
         \centering
         \includegraphics[width=\textwidth]{scatter_plot_500.pdf}
         \caption{$n = 500$}
         \label{subfig:var-sp_500}
     \end{subfigure}
     \begin{subfigure}[h]{0.4\textwidth}
         \centering
         \includegraphics[width=\textwidth]{scatter_plot_50000.pdf}
         \caption{$n = 5 \times 10^4$}
         \label{subfig:var-sp_50000}
     \end{subfigure}
    \caption{Scatter plot of five Markov chains for different sample sizes.}
    \label{fig:var-sp}
\end{figure}

@@

We will use a bivariate bi-modal distribution introduced by \cite{gelman1991note} which has Gaussian conditional distributions in both directions. This allows us to sample parallel Markov chains using the Gibbs sampler. Let $x$ and $y$ be two random variable that are jointly distributed as 
%
\[
f(x, y) \propto \exp\left(-\dfrac{1}{2}[Ax^2y^2 + x^2 + y^2 -2Bxy  -2C_1x - 2C_2y]\right)
\]

The conditional distribution of $x$ given $y$ and vice versa is a normal distribution given by:
%
\begin{align*}
    x_1 \mid x_2 &\sim N\left(\dfrac{Bx_2 + C_1}{Ax_2^2 + 1}, \dfrac{1}{Ax_2^2 + 1}\right)\\
    x_2 \mid x_1 &\sim N\left(\dfrac{Bx_1 + C_2}{Ax_1^2 + 1}, \dfrac{1}{Ax_1^2 + 1}\right)
\end{align*}

@@

In Figure \ref{fig:boom-sp}, we demonstrate the ``sticky'' nature of the Markov chains. For the first thousand samples, both the chains are oblivious of the existence of another mode. By ten thousand samples, both the Markov chains have explored the two modes. This property will affect the single chain estimators like ACF. In Figure \ref{subfig:boom-acf_1e3}, the ACF is severely underestimated because the chain-1 has not jumped to the other mode. Whereas, in Figure \ref{subfig:boom-acf_1e4}, both ACF and R-ACF give almost similar results indicating that both modes have been explored by chain-1.

@@

\begin{table}[h]
\centering
    \begin{tabular}{p{1cm}|p{2cm}|p{2cm}|p{2cm}|p{2cm}}
    % \toprule
        n & \multicolumn{2}{|c|}{$m = 2$} & \multicolumn{2}{|c|}{$m=5$}\\
        \hline
        & ASV & RSV & ASV & RSV\\
        \hline
        $10^3$ & 0.612 & 0.689 & 0.602 & 0.753   \\
        $2\cdot10^3$ & 0.693 & 0.751 & 0.735 & 0.827 \\
        $5 \cdot 10^3$  & 0.826 & 0.854 & 0.847 & 0.880 \\
        $10^4$  & 0.862 & 0.868 & 0.884 & 0.907 \\
        $2 \cdot 10^4$ & 0.899 & 0.906 & 0.922 & 0.934  \\
    % \bottomrule
    \end{tabular}
    \caption{Coverage probabilities for parameter values $A = 1,\; B = 3,\; C = 8$. }
    \label{table:boom-coverage_1_3_8}
\end{table}

@@

\begin{lemma} \label{lemma:consis_1}
  Let the strong invariance principle hold. If $\dfrac{\psi(n)}{n} \to 0$ as $n \to \infty$, then $\| \bar{\bar{X}} - \mu\|_{\infty}, \|\bar{X}_s - \mu\|_{\infty} \xrightarrow[]{a.s.} 0$ as $n \to \infty$ for all $s \in \{1,..., m\}$\\
  \end{lemma}
